---
title: "nr2e3_cluster_anotation_ALL_tp"
author: "Nate Mullin"
date: "9/30/2022"
output: html_document
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(Seurat)
library(cowplot)
library(dplyr)
library(tidyverse)
library(EnhancedVolcano)
library(RColorBrewer)
library(ggplot2)
library(ggpubr)
library(scCustomize)
library(ComplexHeatmap)

set.seed(1234)

```

#################################################################################
###################### LOAD CCA OBJECT AND ADD METADATA #########################
#################################################################################

```{r fig.width=20, fig.height=8}
# load the object using 25 PCA dimensions for UMAP and with stringent filtering
# this is D40, D80, D120, D160, D260 CCA'd together with 25 dims and *stringent* qc cutoffs
load(".../cca_all_tp_10092022/final_object.RData")

# rename object
organoid_all_tp <- IVR_object.combined

# break out metadata
timepoint <- ifelse(
  str_detect(
    organoid_all_tp[["orig.ident"]][["orig.ident"]], "D0"
    ), "D0", ifelse(
  str_detect(
    organoid_all_tp[["orig.ident"]][["orig.ident"]], "D40"
    ), "D40", ifelse(
  str_detect(
    organoid_all_tp[["orig.ident"]][["orig.ident"]], "D80"
    ), "D80", ifelse(
   str_detect(
    organoid_all_tp[["orig.ident"]][["orig.ident"]], "D120"
    ), "D120", ifelse(
  str_detect(
    organoid_all_tp[["orig.ident"]][["orig.ident"]], "D160"
    ), "D160", "D260")))))

line <- ifelse(
  str_detect(
    organoid_all_tp[["orig.ident"]][["orig.ident"]], "B342cor"
    ), "B342cor", ifelse(
  str_detect(
    organoid_all_tp[["orig.ident"]][["orig.ident"]], "B1427"
    ), "B1427", "B342"))

organoid_all_tp[["timepoint"]] <- timepoint
organoid_all_tp[["line"]] <- line


# re-order factors
organoid_all_tp$line <- factor(organoid_all_tp$line, levels = c("B1427", "B342", "B342cor")) # order Rob suggested - WT first
organoid_all_tp$timepoint <- factor(organoid_all_tp$timepoint, levels = c("D40", "D80", "D120", "D160", "D260")) # put in obvious order

table(organoid_all_tp[["timepoint"]])

# save object with metadata
save(organoid_all_tp, file = "organoid_all_tp.RData")

# print the unannotated UMAP split by line and timepoint
pdf("unannotated_umaps.pdf", width=15, height=5)
combo <- DimPlot(organoid_all_tp, shuffle = T, pt.size = 0.05)
time<-DimPlot(organoid_all_tp, group.by = "timepoint", shuffle = T, pt.size = 0.05)
line<-DimPlot(organoid_all_tp, group.by = "line", shuffle = T, pt.size = 0.05)
ggarrange(combo, time, line, nrow = 1, ncol = 3)
dev.off()
```

#################################################################################
############################### Reference Mapping ###############################
#################################################################################

```{r}
# sridhar fetal
fetal_cca_59_82cP_125cp_res1.5_Glia <- readRDS(".../sridhar_data/uw_objects/fetal_cca_59_82cP_125cp_res1.5_Glia.RDS")

#data from Voigt 2021
load(".../2020_09_10_bullseye_manual_analysis.RData")
bullseye <- IVR_object.combined

# unselected donor choroid from Voigt 2020
load(".../2020_06_04_rpechor_exp1_unselected_cca_manual_analysis.RData")
unselected_rpecho <- IVR_object.combined

```

Merge all the references together
```{r fig.width=15, fig.height=5}
DefaultAssay(organoid_all_tp) <- "RNA"

# first combine voigt 2021 and sridhar
merged_ref_data <- merge(bullseye, y = fetal_cca_59_82cP_125cp_res1.5_Glia, add.cell.ids = c("bullseye", "fetal"), project = "merged_ref_data")

# fix metadata
# basically, if it celltype is not NA, use that, otherwise pull the sridhar "type" column
# table(ifelse(!is.na(merged_ref_data$celltype), merged_ref_data$celltype, merged_ref_data$type))
merged_ref_data$celltype <- ifelse(!is.na(merged_ref_data$celltype), 
                                          merged_ref_data$celltype, merged_ref_data$type)

# pull only RPE from unselected - it's 339 cells...
rpe_only <- subset(IVR_object.combined, celltype == "RPE")
table(fetal_cca_59_82cP_125cp_res1.5_Glia$type)

# add RPE to merged reference
merged_ref_data <- merge(merged_ref_data, y = rpe_only, add.cell.ids = c("merged", "rpe"), project = "merged_ref_data")

# okay, there are repeated annotations between voigt and sridhar, will condense.
Idents(merged_ref_data) <- "celltype"
levels(merged_ref_data)

new.cluster.ids <-c(
"Cone"            , "RGC"          ,    "Glia"          , "Cone_ON_BC"    ,   "Cone_OFF_BC"  ,   
"Amacrine"        , "Pericyte"     ,    "Endothelial"     , "Horizontal"    ,   "Rod"          ,   
"Astrocyte"       , "Rod_BC"       ,    "Cone_OFF-BC-BC3A", "Microglia"     ,   "Progenitors"  ,   
"PRC_Progenitor"    , "Amacrine"           ,    "T1_RGC"          , "Early Glia"    ,   "Horizontal"  ,   
"Unknown"             , "T2"           ,    "Glia"            , "T1"            ,   "imGlia"       ,   
"Photoreceptor"  , "T3"           ,    "Bipolars"        , "T2_T3"         ,   "RPE"   )

names(new.cluster.ids) <- levels(merged_ref_data)
merged_ref_data <- RenameIdents(merged_ref_data, new.cluster.ids)
merged_ref_data$celltype <-  Idents(merged_ref_data)

# save reference object
save(merged_ref_data, file = "merged_ref_data.RData")

# finally... kick out donor cell types that are not found in organoids (since RO are pure neuroectoderm.. i think)
`%ni%` <- Negate(`%in%`)
merged_ref_data <- subset(merged_ref_data, celltype %ni% c("Pericyte", "Endothelial", "Microglia", "Unknown", "Photoreceptor", "PRC_Progenitor"))

# now we can do the label transfer!
DefaultAssay(organoid_all_tp) <- "RNA"
DefaultAssay(merged_ref_data) <- "RNA"

common.features <- intersect(rownames(organoid_all_tp), rownames(merged_ref_data))
length(x = common.features)

# ugh ned to do this...
merged_ref_data <- FindVariableFeatures(merged_ref_data)
merged_ref_data <- ScaleData(merged_ref_data)
merged_ref_data <- RunPCA(merged_ref_data)

# find anchors between the fetal and organoid data
anchors <- FindTransferAnchors(reference = merged_ref_data, 
                               query = organoid_all_tp,
                               dims = 1:30, reference.reduction = "pca",
                               features = common.features, reduction = "pcaproject")


# derive cell type predictions using anchors. Use Sridhar's annotations which include T1, T2, etc.
predictions <- TransferData(anchorset = anchors, refdata = merged_ref_data$celltype,
    dims = 1:30)

# add predictions to organoid object and label as predicted.id
organoid_all_tp <- AddMetaData(organoid_all_tp, metadata = predictions)
Idents(organoid_all_tp) <- 'predicted.id'

# print a plot showing prediction score
pdf("merged_wRPE_pred_score_stringent.pdf", width=5, height=5)
FeaturePlot(organoid_all_tp, features =  "prediction.score.max")
dev.off()

# print a vln plot showing prediction score
pdf("merged_wRPE_pred_score_stringent_vln.pdf", width=10, height=5)
VlnPlot(organoid_all_tp, features =  "prediction.score.max")
dev.off()

# so-called microglia are low prediction score which makes sense - drop everything under 0.25 to be safe.
organoid_all_tp$donor_annotation <- ifelse(organoid_all_tp$prediction.score.max > 0.25, 
                                                 organoid_all_tp$predicted.id, NA)

Idents(organoid_all_tp) <- 'donor_annotation'

#print UMAP
pdf("dimplot_donor_annotation.pdf", width=5, height=5)
DimPlot(organoid_all_tp, reduction = "umap", shuffle = T, label = T, group.by = "donor_annotation", ncol = 1) + NoLegend() + NoAxes()
dev.off()

# drop NA's
organoid_all_tp <- subset(organoid_all_tp, donor_annotation != "NA")
```

##################################################################################
############################### REFINE ANNOTATIONS ###############################
##################################################################################

An issue above is that Transition populations (by Sridhar) are being lost/misannotated by Voigt data... will try to fix manually by subclustering?

PLAN: Merge T1,T3 -> then break out 
```{r}
# it looks like all Sridhar populations are jammed into T1 - try subclustering to find T2 and T3.
DimPlot(subset(organoid_all_tp, donor_annotation == "T1"), label = T, split.by = "timepoint")


levels(organoid_all_tp)

new.cluster.ids <-c(
"Glia"        ,     "T_PRC"         ,     "T_PRC"      ,       "Cone_OFF_BC"     , "Cone_ON_BC"  ,    
"Amacrine"    ,     "Rod_BC"      ,     "Astrocyte" ,       "T2"              , "Early Glia"  ,    
"RGC"         ,     "RPE"         ,     "T_PRC"        ,       "Cone_OFF-BC-BC3A", "Horizontal"  ,    
"Progenitors" ,     "imGlia"      ,     "T_PRC"  )

names(new.cluster.ids) <- levels(organoid_all_tp)
organoid_all_tp <- RenameIdents(organoid_all_tp, new.cluster.ids)
organoid_all_tp$donor_annotation <-  Idents(organoid_all_tp)

# not sure what happened to old graph.. ?
organoid_all_tp <- FindNeighbors(organoid_all_tp, reduction = "pca", dims = 1:25)

Idents(organoid_all_tp) <- "donor_annotation"
organoid_all_tp <- FindSubCluster(object = organoid_all_tp, cluster = "T_PRC", graph.name = "integrated_snn", algorithm = 3, resolution = 0.5)
  # try using Leiden.

# look for subclusters

Idents(organoid_all_tp) <- "sub.cluster"
DimPlot(organoid_all_tp, reduction = "umap", label = T)

markers <- rev(c("HES6", "ATOH7", "DLL1", #T1
                     "PRDM13", "HSPB1", #T2
                     "DLL3", "OTX2", "FABP7", #T3
                     "ARR3", "PDE6H",
                      "RHO", "NRL", "NR2E3"
))

pdf(file = "sridhar_t_pop_marker_dot_plot.pdf", width = 15, height = 4)
DotPlot_scCustom(organoid_all_tp, features = markers, flip_axes = T, x_lab_rotate = T)
dev.off()

levels(organoid_all_tp)
table(organoid_all_tp$sub.cluster)

new.cluster.ids <-c(
"Glia"        , "Rod"         ,"Rod"    ,"Cone"  ,"Cone_OFF_BC"  ,   
"Cone"     , "Cone_ON_BC"      ,"Cone"    ,"Amacrine" ,"Rod_BC"       ,   
"Astrocyte"   , "Rod"         ,"T2"         ,"Rod" ,"Early Glia"   ,   
"T3"     , "RGC"             ,"RPE"        ,"Rod" ,"T3"     ,   
"T3"     , "Cone_OFF-BC-BC3A","Horizontal" ,"T3" ,"Progenitors"  ,   
"T3"     , "T3"        ,"T1"    ,"imGlia"   ,"T1"     ,   
"T2"    , "T3"        ,"T1"   ,"T3"   
)

names(new.cluster.ids) <- levels(organoid_all_tp)
organoid_all_tp <- RenameIdents(organoid_all_tp, new.cluster.ids)
organoid_all_tp$donor_annotation <-  Idents(organoid_all_tp)
Idents(organoid_all_tp) <- "donor_annotation"

DimPlot(organoid_all_tp, reduction = "umap", label = T)
DimPlot(subset(organoid_all_tp, donor_annotation == "T3"), label = T)

```

Now do the same thing with the Photoreceptor population
These definitely have some differentiated rods and cones. If impossible to tell will leave at "PRC_Progenitor"
```{r}
# it looks like all Sridhar populations are jammed into T1 - try subclustering to find T2 and T3.
DimPlot(subset(organoid_all_tp, donor_annotation == "T3"), label = T) + NoLegend()

# not sure what happened to old graph.. ?
Idents(organoid_all_tp) <- "donor_annotation"
organoid_all_tp <- FindSubCluster(object = organoid_all_tp, cluster = "Photoreceptor", graph.name = "integrated_nn", algorithm = 1, resolution = 0.05)
  # try using Leiden.

# look for subclusters

Idents(organoid_all_tp) <- "sub.cluster"
DimPlot(organoid_all_tp, reduction = "umap", label = T) + NoLegend()

sridhar_markers <- rev(c("HES6", "ATOH7", "DLL1", #T1
                     "PRDM13", "HSPB1", #T2
                     "DLL3", "OTX2", "FABP7" #T3
))

pdf(file = "sridhar_t_pop_marker_dot_plot.pdf", width = 15, height = 3)
DotPlot_scCustom(organoid_all_tp, features = sridhar_markers, flip_axes = T, x_lab_rotate = T)
dev.off()

levels(organoid_all_tp)

new.cluster.ids <-c(
"Glia"      ,       "Rod"        ,      "Cone"           ,  "Cone_OFF_BC"   ,  "Cone_ON_BC"   ,  
"Rod_BC"    ,       "Astrocyte"  ,      "PRC_Progenitor"  , "Photoreceptor" ,  "Early Glia"   ,  
"Amacrine"  ,       "RPE"        ,      "Horizontal",       "T3" ,            "Progenitors" ,   
"Cone_OFF-BC-BC3A", "RGC"      ,        "T1"   ,        "T1"  ,           "T1"            ,
"T3"       ,      "T3"         ,    "T1"       ,     "T2"        ,     "T3"            ,
"T1"       ,     "T2"          ,   "T3"        ,    "T3"         ,   "T3"           ,
"T3"       ,     "T3"          ,  "T3"         ,   "T1"          ,  "T1"           ,
"T1"       ,      "imGlia"     ,      "T1"     ,       "T3"      ,      "T1"           ,
"T3" )

names(new.cluster.ids) <- levels(organoid_all_tp)
organoid_all_tp <- RenameIdents(organoid_all_tp, new.cluster.ids)
organoid_all_tp$donor_annotation <-  Idents(organoid_all_tp)

DimPlot(organoid_all_tp, reduction = "umap", label = T)
```

CLEAN UP T3
```{r}
organoid_all_tp <- FindSubCluster(object = organoid_all_tp, cluster = "T3", graph.name = "integrated_nn", algorithm = 1, resolution = 0.05)
  # try using Leiden.

# look for subclusters

Idents(organoid_all_tp) <- "sub.cluster"
DimPlot(organoid_all_tp, reduction = "umap", label = T) + NoLegend()
table(organoid_all_tp$sub.cluster)

pdf(file = "donor_ann_clusters.pdf", width = 30,height = 30)
DimPlot(organoid_all_tp, label = T, split.by = "sub.cluster", ncol = 5) + NoLegend()
dev.off()


pdf(file = "sridhar_t_pop_marker_dot_plot.pdf", width = 15, height = 3)
DotPlot_scCustom(organoid_all_tp, features = sridhar_markers, flip_axes = T, x_lab_rotate = T)
dev.off()

levels(organoid_all_tp)

new.cluster.ids <-c(
 "Glia"          ,   "Rod"            ,  "Cone"          ,   "Cone_OFF_BC"    ,  "Cone_ON_BC"      ,
 "Amacrine"      ,   "Rod_BC"         ,  "Astrocyte"     ,   "T2"             ,  "Early Glia"      ,
 "Discard"          ,   "RGC"            ,  "RPE"           ,   "T3"           ,  "Cone_OFF-BC-BC3A",
 "Horizontal"    ,   "Progenitors"    ,  "Discard"          ,   "T1"             ,  "Discard"            ,
 "imGlia"        ,   "Discard"          ,  "Discard"          ,   "Discard"           ,  "Discard"            ,
 "Discard"          ,   "Discard"           ,  "Discard"         ,   "Discard"     )

names(new.cluster.ids) <- levels(organoid_all_tp)
organoid_all_tp <- RenameIdents(organoid_all_tp, new.cluster.ids)
organoid_all_tp$donor_annotation <-  Idents(organoid_all_tp)

DimPlot(organoid_all_tp, reduction = "umap", label = T)

```


Reorder celltypes
```{r}
table(organoid_all_tp$donor_annotation)
organoid_all_tp$donor_annotation <- factor(organoid_all_tp$donor_annotation, levels = rev(c("Progenitors", "T1", "T2","T3" ,"RGC" ,"Early Glia", "imGlia", "Glia", "Amacrine", "Astrocyte", "Horizontal", "RPE",  "Cone", "Rod", "Cone_OFF_BC", "Cone_OFF-BC-BC3A", "Cone_ON_BC", "Rod_BC", "Discard")))

mutate(my_cols, factor(organoid_all_tp$donor_annotation))
my_cols
length(levels(organoid_all_tp$donor_annotation))

levels(organoid_all_tp$donor_annotation)
my_cols[1:19]
show_col(my_cols[1:19])

my_new_cols <- c("#666666", "#FB9A99", "#D95F02", "#1B9E77", "#7570B3",
                 "#E31A1C", "#1F78B4", "#33A02C", "#66A61E", "#66C2A5",
                 "#FF7F00", "#6A3D9A", "#E7298A", "#CAB2D6", "#B2DF8A",
                 "#E6AB02", "#A6761D", "#FDBF6F", "#A6CEE3")

rev(my_new_cols)
```



Finally print a table of stats
```{r}
cluster_stats <- Cluster_Stats_All_Samples(seurat_object = organoid_all_tp)
write.csv(cluster_stats, file = "final_annotation_stats.csv", col.names = T, row.names = T)
save(organoid_all_tp, file = "organoid_all_tp.RData")

```


```{r}
table(organoid_all_tp$donor_annotation)

organoid_all_tp <- subset(organoid_all_tp, donor_annotation != "Discard")
save(organoid_all_tp, file = "organoid_all_tp_no_discard.RData")
```

